Language models as probabilistic cognitive models

### 2023-11-06 03-paper/paper-LLM-probmod.tex
master table, rev §4

As a concrete example, we introduce the Rational Speech Act (RSA) model first described in this form by \citet{FrankGoodman2012:Predicting-Prag} \citep[for overview see][]{FrankeJager2015:Probabilistic-p,GoodmanFrank2016:Pragmatic-Langu,StevensBenz2018:Game-Theoretic-,ScontrasTessler2021:A-practical-int,Degen2023:The-Rational-Sp}.
The RSA model defines pragmatic reasoning as a sequence of iterated (soft-)optmization of policies, grounding out in literal interpretation.
If $\mathfrak{L}(s,u) \mapsto \set{0,1}$ is a semantic meaning function mapping each pair of state $s$ and utterance $u$ to a (binary) truth-value, and if $P_{s}$ is a prior over states, a literal listener policy is defined as:
%
\begin{align*}
 P_{L_{0}}(s \mid u) \propto \mathfrak{L}(s,u) P(s)\,.
\end{align*}
%
The pragmatic speaker policy is defined as soft-optimizing the choice of utterance to minimize the literal listener's surprisal for the state to be communicated, i.e., the trigger object:
%
\begin{align*}
  P_{S}(u_{i} \mid s, \alpha) & \propto \expo \left [ \log P_{L_{0}}(s \mid u_{i}) \right ] \,.
\end{align*}
%
Finally, the pragmatic listener is defined as the policy resulting from applying Bayes rule, solving the inverse-problem for the previously defined speaker policy:
%
\begin{align*}
  P_{L}(s \mid u, \alpha) \propto  P_{S}(u \mid s, \alpha) \  P(s) \,.
\end{align*}

@@

Figure~\ref{fig:RSA-example} gives example calculations (assuming a flat prior $P(s)$ and $\alpha=1$) for the reference game from Figure~\ref{fig:ref-game}.
For $\alpha=1$, the model predicts that the probabilities of target, competitor and distractor options are $\tuple{\nicefrac{2}{3}, \nicefrac{1}{3}, 0}$ for the production, and $\tuple{\nicefrac{3}{5}, \nicefrac{2}{5}, 0}$ for the interpretation condition.
Increasing $\alpha$ will increase the odds of target over competitor choices.
Yet, the model also predicts probability zero for distractor choices, so that the human data shown in Figure~\ref{fig:refgame-counts}, where the distractor option was chosen in both conditions, would immediately rule out the model entirely.
It is therefore common to include a small error probability $\epsilon$, with which a choice would be made at random \citep[e.g.,][]{LeeWagenmakers2013:Bayesian-Cognit}: if $P_{r}(R_{l}, C, \alpha_{c})$ is the RSA model's probabilistic prediction for response category $R_{l}$ (target, competitor, or distractor) for condition $C$ (production or interpretation) and condition-specific optimality $\alpha_{c}$, the random-error smoothed prediction is:
\begin{align*}
  P_{r}(R_{l}, C; \alpha_{c}, \epsilon_{c}) = (1 - \epsilon) \  P_{r}(R_{l}, C; \alpha_{c}) +  \nicefrac{\epsilon_{c}}{3}\,,
\end{align*}
where $\epsilon_{c}$ is a (condition-specific) parameter giving the probability that a choice was made by randomly guessing.
\mf{FC objected to ``condition-specificity'' which is fair, but I leave it as it was; my reasoning: we use model checking to rule out models; in this setting the conservative strategy is to err on the side of making the models too powerful (even at the cost of conceptual plausibility)}
The data $D_{C}$ from condition $C$, see Figure~\ref{fig:refgame-counts}, consists of counts for each response category.
The parameterized likelihood function entailed by the RSA model for $D_{C}$ is:
%
\begin{align*}
 P^{\text{RSA}}_{\text{cond}}(D_{C} \mid C, \alpha_{C}, \epsilon_{C}) = \text{Multinomial}(D_{C}, \tuple{P_{r}(R_{l}, C; \alpha_{c}, \epsilon_{c}}_{1 \le l \le 3})\,.
\end{align*}
%
The result is a four-parameter model, one pair of parameters per condition.

Parameterized predictions like $P_{r}(D_{C} \mid C, \alpha_{C}, \epsilon_{C})$ can be assessed in the light of the empirical data with the usual tools of Bayesian data analysis \citep[e.g.][]{GelmanCarlin2014:Bayesian-Data-A,McElreath2016:Statistical-Ret,Lambert2018:A-Students-Guid}.
Let $\alpha_{c}\sim \text{log-Normal}(0.5,1)$ have a reasonably wide log-Normal prior, and let $\epsilon_{c} \sim \text{Beta}(1,5)$ have a Beta prior favoring small values.
Using Stan \citep{Team2023:The-Stan-Core-L} for Bayesian inference, we obtain estimates of posterior credible values of model parameters (summary statistics of which are shown in Table~\ref{fig:posterior-stats}).\footnote{
  Posterior samples where generated for four chains with 2000 samples each, after a warm-up of 1000 samples. The ``adapt-delta'' value was set to 0.99. Converge was checked with $\hat{R}$-statistics \citep{GelmanRubin1992:Inference-from-}.
}

To assess goodness-of-fit, we use the \emph{posterior predictive distribution}, i.e., the model's predictions about data of the same size and structure as the training data.
Figure~\ref{fig:refgame-counts} shows summary statistics (means and 95\% credible intervals) for the posterior predictive distribution of the RSA model (among other models, which are to be introduced later).
We see that for both conditions the RSA model passes a ``visual posterior predictive check'' \citep{Kruschke2011:Doing-Bayesian-}, namely that the distribution of posterior predictions includes the observed choice rates for each answer option.
To corroborate the visual impression, Table~\ref{tab:Bppp-values} shows sample-based estimates of Bayesian posterior predictive $p$-values, using likelihood of the observed data as a test statistics.
These values approximate the probability that a model trained on $D_{\text{obs}}$ would predict future data of the same size and format that is at least as unlikely as the data $D_{\text{obs}}$ itself.
A very small value indicates that the model might be inadequate for reproducing the data it was trained on.
This is clearly not the case for the RSA model with posterior predictive $p$-values close to 0.5.
In sum, the condition-level predictions by the RSA model, a theoretically motivated PCM, are not discredited by the condition-level data.

We can also use the condition-level predictions for predicting the item-level data.
If the whole data set for condition $C$ is chunked into item-level data, $D_{C} = \tuple{D^{1}_{C}, \dots, D^{m}_{C} }$, the likelihood function for analysis at the item-level uses the same (condition-level) predictor to separately predict data from each item:
%
\begin{align*}
 P^{\text{RSA}}_{\text{item}}(D_{C} \mid C, \alpha_{C}, \epsilon_{C}) = \prod_{k = 1}^{m} \text{Multinomial}(D_{C}^{k}, \tuple{P_{r}(R_{l}, C; \alpha_{c}, \epsilon_{c}}_{1 \le l \le 3})\,.
\end{align*}
%
Fitting for the partitioned data set, and estimates of posterior predictive $p$-values that do not discredit the model ($0.338$ for production, $0.252$ for interpretation).
This suggests that the item-level variation in the human data is not so pronounced as to provide strong evidence against the condition-level predictor from a theoretically motivated PCM.

@@

@@

\begin{table}[ht]
\centering
\begin{tabular}{rlrrrrrr}
  \toprule
 && $\alpha$ &&& \(\epsilon\) & \\
 & condition & $|$95 & mean & 95\%$|$ & $|$95\% & mean\_epsilon & 95\%$|$\_epsilon \\ \midrule
1 & production & 0.15 & 0.21 & 0.29 & 0.00 & 0.10 & 0.21 \\
  2 & interpretation & 0.08 & 0.13 & 0.18 & 0.00 & 0.17 & 0.33 \\
   \bottomrule
\end{tabular}
\end{table}

\mf{Posteriors over $\alpha$ and $\epsilon$}

@@

\begin{table}[t]
  \centering

  % \includegraphics[width=0.9\linewidth]{00-pics/posterior-stats.pdf}

  \begin{tabular}{llllcr}
    \toprule \addlinespace[1ex]
    model        & condition      & Parameter  & $|$95\% & mean & 95\%$|$ \\
    \midrule  \addlinespace[1ex]
    narrow       & production     & $\alpha$   & 0.19    & 0.22 & 0.26 \\
                 &                & $\epsilon$ & 0.00    & 0.04 & 0.11 \\
                 & interpretation & $\alpha$   & 0.15    & 0.18 & 0.20 \\
                 &                & $\epsilon$ & 0.00    & 0.02 & 0.06 \\ \addlinespace[0.75ex]
    intermediate & production     & $\alpha$   & 0.86    & 1.04 & 1.22 \\
                 &                & $\epsilon$ & 0.08    & 0.13 & 0.18 \\
                 & interpretation & $\alpha$   & 0.36    & 0.48 & 0.60 \\
                 &                & $\epsilon$ & 0.00    & 0.05 & 0.12 \\ \addlinespace[0.75ex]
    wide         & production     & $\alpha$   & 0.22    & 2.69 & 8.79 \\
                 &                & $\epsilon$ & 0.08    & 0.13 & 0.19 \\
                 & interpretation & $\alpha$   & 0.19    & 1.08 & 6.03 \\
                 &                & $\epsilon$ & 0.00    & 0.06 & 0.18 \\ \addlinespace[0.75ex]
    RSA          & production     & $\alpha$   & 2.59    & 3.16 & 3.71 \\
                 &                & $\epsilon$ & 0.08    & 0.13 & 0.18 \\
                 & interpretation & $\alpha$   & 0.24    & 0.66 & 1.07 \\
                 &                & $\epsilon$ & 0.10    & 0.15 & 0.20 \\ \addlinespace[0.25ex]
    \bottomrule \\
  \end{tabular}

  \caption{
    Summary statistics of posterior samples.
    Numbers indicate estimates for lower and upper bounds of 95\% credible intervals and the posterior means.
    \mf{make wide format}
  }
  \label{fig:posterior-stats}
\end{table}

The three condition-level predictions yield three different probabilistic models, each with two pairs of condition-dependent parameters $\alpha_{c}$, and $\epsilon_{c}$.
Using priors and Bayesian posterior estimation as previously described for the RSA (see Section~\ref{sec:model-pred-from}), we obtain samples from the posterior over parameters and samples from the posterior predictive distributions.
The usual Bayesian summary statistics for posteriors over model parameters are shown in Table~\ref{fig:posterior-stats}.
What is noteworthy about these estimates is that the $\alpha_{c}$ parameters for the wide-scope model have a very large credible range.
This is because, already for values of $\alpha_{c} = 1$, the item-level predictions are virtually ``winner-takes-all'' choices, selecting the option with the maximum score with probability almost 1 for almost all items.
This implies that, by studying the posterior predictive distribution of the wide-scope model, we are virtually testing condition-level predictions from the standard WTA-strategy used for benchmark testing (see Section~\ref{motivation}).

Figure~\ref{fig:refgame-counts} shows the summary statistics (means and 95\% credible intervals) for each model's posterior predictive distribution.
We find that only the theoretical model (RSA) and the intermediate-scope model passes this ``visual posterior predictive check'' for both conditions; the other two models both overpredict the target choice rate and underpredict the competitor choice rate in the interpretation condition.
To corroborate the visual impression, Table~\ref{tab:Bppp-values} shows sample-based estimates of Bayesian posterior predictive $p$-values, using likelihood of the observed data as a test statistics.
Consequently, the results from Table~\ref{tab:Bppp-values} suggest that the narrow-scope model fails on the interpretation data, and is at most borderline compatible with the production data; that the wide-scope model is able to reproduce the production data, but fails on the interpretation data; and that only the intermediate-scope model is able to fully recover the data from both conditions.

\begin{table}[ht]
\centering

% \begin{tabular}{lrr}
%   \toprule
%   model        & production & interpretation \\ \midrule
%   narrow       & 0.05       & 0.00 \\
%   wide         & 0.51       & 0.00 \\
%   intermediate & 0.49       & 0.26 \\
%   RSA          & 0.49       & 0.52 \\
%   \bottomrule \\
% \end{tabular}

\begin{tabular}{lrrrr}
  \toprule
  & narrow & wide & intermediate & RSA \\ \midrule
  production & 0.05 & 0.51 & 0.49 & 0.49 \\
  interpretation & 0.00 & 0.00 & 0.26 & 0.52 \\ \bottomrule \\
\end{tabular}

\caption{Sample-based estimates of Bayesian posterior predictive $p$-values for each model and each condition, based on likelihood of the observed data as test statistic.}
\label{tab:Bppp-values}
\end{table}

>>>>

As a concrete example, we introduce the Rational Speech Act (RSA) model first described in this form by \citet{FrankGoodman2012:Predicting-Prag} \citep[for overview see][]{FrankeJager2015:Probabilistic-p,GoodmanFrank2016:Pragmatic-Langu,StevensBenz2018:Game-Theoretic-,ScontrasTessler2021:A-practical-int,Degen2023:The-Rational-Sp}.
The RSA model defines pragmatic reasoning as a sequence of iterated (soft-)optmization of policies, grounding out in literal interpretation.
If $\mathfrak{L}(s,u) \mapsto \set{0,1}$ is a semantic meaning function mapping each pair of state $s$ and utterance $u$ to a (binary) truth-value, and if $P_{\text{prior}}(s)$ is a prior over states, a literal listener policy is defined as:
%
\begin{align*}
 P_{L_{0}}(s \mid u) \propto \mathfrak{L}(s,u)\  P_{\text{prior}}(s)\,.
\end{align*}
%
The pragmatic speaker policy is defined as soft-optimizing the choice of utterance to minimize the literal listener's surprisal for the state to be communicated, i.e., to maximize the log-probability of the trigger object:
%
\begin{align}
  \label{eq:RSA-speaker}
  P_{S}(u_{i} \mid s, \alpha) & \propto \expo \left [ \log P_{L_{0}}(s \mid u_{i}) \right ] \,.
\end{align}
%
Finally, the pragmatic listener is defined as the policy resulting from applying Bayes rule, solving the inverse-problem for the previously defined speaker policy:
%
\begin{align}
  \label{eq:RSA-interpreter}
  P_{L}(s \mid u, \alpha) \propto  P_{S}(u \mid s, \alpha) \  P_{\text{prior}}(s) \,.
\end{align}

@@

Figure~\ref{fig:RSA-example} gives example calculations (assuming a flat prior and $\alpha=1$) for the reference game from Figure~\ref{fig:ref-game}.
For $\alpha=1$, the model predicts that the probabilities of target, competitor and distractor options are $\tuple{\nicefrac{2}{3}, \nicefrac{1}{3}, 0}$ for the production, and $\tuple{\nicefrac{3}{5}, \nicefrac{2}{5}, 0}$ for the interpretation condition.
Increasing $\alpha$ will increase the odds of target over competitor choices.
In sum, the condition-level predictions of the RSA model are a parameterized function $P_{\text{cond}}^{\text{RSA}}(R_{l}, C; \alpha_{c})$, assigning a probability to each response category $R_{l}$ in each condition $C$ (production or interpretation).

The model constructed so far predicts probability zero for distractor choices, so that the human data shown in Figure~\ref{fig:refgame-counts}, where the distractor option was chosen in both conditions, would immediately rule out the model entirely.
It is therefore common to include a small error probability $\epsilon$, with which a choice would be made at random \citep[e.g.,][]{LeeWagenmakers2013:Bayesian-Cognit}, so that we get:
%
\begin{align*}
  P_{r}(R_{l}, C; \alpha_{c}, \epsilon_{c}) = (1 - \epsilon) \  P_{r}(R_{l}, C; \alpha_{c}) +  \nicefrac{\epsilon_{c}}{3}\,,
\end{align*}
%
where $\epsilon_{c}$ is a (condition-specific) parameter giving the probability that a choice was made by randomly guessing.
\mf{FC objected to ``condition-specificity'' which is fair, but I leave it as it was; my reasoning: we use model checking to rule out models; in this setting the conservative strategy is to err on the side of making the models too powerful (even at the cost of conceptual plausibility)}

The data $D_{C}$ from condition $C$, see Figure~\ref{fig:refgame-counts}, consists of counts for each response category.
The parameterized likelihood function entailed by the RSA model for condition-level data $D_{C}$ is:
%
\begin{align}
  \label{eq:RSA-likelihood}
 P^{\text{RSA}}_{\text{cond}}(D_{C} \mid C, \alpha_{C}, \epsilon_{C}) = \text{Multinomial}(D_{C}, \tuple{P_{r}(R_{l}, C; \alpha_{c}, \epsilon_{c}}_{1 \le l \le 3})\,.
\end{align}
%
The result is a four-parameter model, one pair of parameters per condition.

Parameterized predictions, like in Equation~(\ref{eq:RSA-likelihood}), can be assessed in the light of the empirical data with the usual tools of Bayesian data analysis \citep[e.g.][]{GelmanCarlin2014:Bayesian-Data-A,McElreath2016:Statistical-Ret,Lambert2018:A-Students-Guid}.
Let $\alpha_{c}\sim \text{log-Normal}(0.5,1)$ have a reasonably wide log-Normal prior, and let $\epsilon_{c} \sim \text{Beta}(1,5)$ have a Beta prior favoring small values.
Using Stan \citep{Team2023:The-Stan-Core-L} for Bayesian inference, we obtain estimates of posterior credible values of model parameters (summary statistics of which are shown in Table~\ref{tab:sumStats}).\footnote{
  Posterior samples where generated for four chains with 2000 samples each, after a warm-up of 1000 samples. The ``adapt-delta'' value was set to 0.99. Converge was checked with $\hat{R}$-statistics \citep{GelmanRubin1992:Inference-from-}.
}

To assess goodness-of-fit, we use the \emph{posterior predictive distribution}, i.e., the model's predictions about data of the same size and structure as the training data.
Figure~\ref{fig:refgame-counts} shows summary statistics (means and 95\% credible intervals) for the posterior predictive distribution of the RSA model (among other models, which are to be introduced later).
We see that for both conditions the RSA model passes a ``visual posterior predictive check'' \citep{Kruschke2011:Doing-Bayesian-}, namely that the distribution of posterior predictions includes the observed choice rates for each answer option.
To corroborate the visual impression, Table~\ref{tab:sumStats} shows sample-based estimates of Bayesian posterior predictive $p$-values, using likelihood of the observed data as a test statistics.
These values approximate the probability that a model trained on $D_{\text{obs}}$ would predict future data of the same size and format that is at least as unlikely as the data $D_{\text{obs}}$ itself.
A very small value indicates that the model might be inadequate for reproducing the data it was trained on.
This is clearly not the case for the RSA model with posterior predictive $p$-values close to 0.5, as shown in Table~\ref{tab:sumStats}.
In sum, the condition-level predictions by the RSA model, a theoretically motivated PCM, are not discredited by the condition-level data.

We can also use the condition-level predictions for predicting the item-level data.
If the whole data set for condition $C$ is chunked into item-level data, $D_{C} = \tuple{D^{1}_{C}, \dots, D^{m}_{C} }$, the likelihood function for analysis at the item-level uses the same (condition-level) predictor to separately predict data from each item:
%
\begin{align*}
 P^{\text{RSA}}_{\text{item}}(D_{C} \mid C, \alpha_{C}, \epsilon_{C}) = \prod_{k = 1}^{m} \text{Multinomial}(D_{C}^{k}, \tuple{P_{r}(R_{l}, C; \alpha_{c}, \epsilon_{c}}_{1 \le l \le 3})\,.
\end{align*}
%
Fitting for the partitioned data set, and estimates of posterior predictive $p$-values that do not discredit the model (see Table~\ref{tab:sumStats}).
This suggests that the item-level variation in the human data is not so pronounced as to provide strong evidence against the condition-level predictor from a theoretically motivated PCM.

@@

\begin{table}[t]
\centering
\begin{tabular}{llllcrlcrcl}
  \toprule
  &&&& $\alpha$ &&& $\epsilon$ & \\ \cmidrule(r){4-6} \cmidrule(l){7-9}
  data & condition & model & $|$95\%\ & mean\ & 95\%$|$\ & $|$95\%\ & mean\ & 95\%$|$\ & Bpppv & \\
  \midrule

  cond.-level & avg. scores        & production     & 0.19 & 0.00 & 0.22 & 0.04 & 0.26 & 0.11 & 0.05 & * \\
  cond.-level & avg. scores        & interpretation & 0.15 & 0.00 & 0.18 & 0.02 & 0.20 & 0.06 & 0.00 & * \\
  cond.-level & avg. probabilities & production     & 0.86 & 0.08 & 1.04 & 0.13 & 1.22 & 0.18 & 0.51 &  \\
  cond.-level & avg. probabilities & interpretation & 0.36 & 0.00 & 0.48 & 0.05 & 0.60 & 0.12 & 0.00 & * \\
  cond.-level & avg. WTA           & production     & 0.87 & 0.08 & 1.04 & 0.13 & 1.23 & 0.17 & 0.49 &  \\
  cond.-level & avg. WTA           & interpretation & 0.17 & 0.10 & 0.39 & 0.15 & 0.61 & 0.20 & 0.26 &  \\
  cond.-level & RSA                & production     & 2.59 & 0.08 & 3.16 & 0.13 & 3.71 & 0.18 & 0.49 &  \\
  cond.-level & RSA                & interpretation & 0.24 & 0.10 & 0.66 & 0.15 & 1.07 & 0.20 & 0.52 &  \\
  item-level  & LLM                & production     & 0.15 & 0.00 & 0.21 & 0.10 & 0.29 & 0.21 & 0.00 & * \\
  item-level  & LLM                & interpretation & 0.08 & 0.00 & 0.13 & 0.17 & 0.18 & 0.33 & 0.00 & * \\
  item-level  & RSA                & production     & 2.62 & 0.08 & 3.15 & 0.13 & 3.69 & 0.17 & 0.34 &  \\
  item-level  & RSA                & interpretation & 0.24 & 0.10 & 0.66 & 0.15 & 1.05 & 0.20 & 0.25 & \\

   \bottomrule \\
\end{tabular}
\caption{
  Summary statistics for model fits.
  Columns show the means and 95\% credible intervals for posterior samples for each parameter, as well as the Bayesian posterior predictive $p$-value for each model, condition and data level.
  Asterisks mark $p$-values not above $0.05$, flagging relative inability to capture the structure of the data.
}
\label{tab:sumStats}
\end{table}

@@

@@

% \begin{table}[t]
%   \centering

%   % \includegraphics[width=0.9\linewidth]{00-pics/posterior-stats.pdf}

%   \begin{tabular}{llllcr}
%     \toprule \addlinespace[1ex]
%     model        & condition      & Parameter  & $|$95\% & mean & 95\%$|$ \\
%     \midrule  \addlinespace[1ex]
%     narrow       & production     & $\alpha$   & 0.19    & 0.22 & 0.26 \\
%                  &                & $\epsilon$ & 0.00    & 0.04 & 0.11 \\
%                  & interpretation & $\alpha$   & 0.15    & 0.18 & 0.20 \\
%                  &                & $\epsilon$ & 0.00    & 0.02 & 0.06 \\ \addlinespace[0.75ex]
%     intermediate & production     & $\alpha$   & 0.86    & 1.04 & 1.22 \\
%                  &                & $\epsilon$ & 0.08    & 0.13 & 0.18 \\
%                  & interpretation & $\alpha$   & 0.36    & 0.48 & 0.60 \\
%                  &                & $\epsilon$ & 0.00    & 0.05 & 0.12 \\ \addlinespace[0.75ex]
%     wide         & production     & $\alpha$   & 0.22    & 2.69 & 8.79 \\
%                  &                & $\epsilon$ & 0.08    & 0.13 & 0.19 \\
%                  & interpretation & $\alpha$   & 0.19    & 1.08 & 6.03 \\
%                  &                & $\epsilon$ & 0.00    & 0.06 & 0.18 \\ \addlinespace[0.75ex]
%     RSA          & production     & $\alpha$   & 2.59    & 3.16 & 3.71 \\
%                  &                & $\epsilon$ & 0.08    & 0.13 & 0.18 \\
%                  & interpretation & $\alpha$   & 0.24    & 0.66 & 1.07 \\
%                  &                & $\epsilon$ & 0.10    & 0.15 & 0.20 \\ \addlinespace[0.25ex]
%     \bottomrule \\
%   \end{tabular}

%   \caption{
%     Summary statistics of posterior samples.
%     Numbers indicate estimates for lower and upper bounds of 95\% credible intervals and the posterior means.
%     \mf{make wide format}
%   }
%   \label{fig:posterior-stats}
% \end{table}

The three condition-level predictions yield three different probabilistic models, each with two pairs of condition-dependent parameters $\alpha_{c}$, and $\epsilon_{c}$.
Using priors and Bayesian posterior estimation as previously described for the RSA (see Section~\ref{sec:model-pred-from}), we obtain samples from the posterior over parameters and samples from the posterior predictive distributions.
The usual Bayesian summary statistics for posteriors over model parameters are shown in Table~\ref{tab:sumStats}.
What is noteworthy about these estimates is that the $\alpha_{c}$ parameters for the wide-scope model have a very large credible range.
This is because, already for values of $\alpha_{c} = 1$, the item-level predictions are virtually ``winner-takes-all'' choices, selecting the option with the maximum score with probability almost 1 for almost all items.
This implies that, by studying the posterior predictive distribution of the wide-scope model, we are virtually testing condition-level predictions from the standard WTA-strategy used for benchmark testing (see Section~\ref{motivation}).

Figure~\ref{fig:refgame-counts} shows the summary statistics (means and 95\% credible intervals) for each model's posterior predictive distribution.
We find that only the theoretical model (RSA) and the intermediate-scope model passes this ``visual posterior predictive check'' for both conditions; the other two models both overpredict the target choice rate and underpredict the competitor choice rate in the interpretation condition.
To corroborate the visual impression, Table~\ref{tab:sumStats} shows sample-based estimates of Bayesian posterior predictive $p$-values, using likelihood of the observed data as a test statistics.
Consequently, the results from Table~\ref{tab:sumStats} suggest that the narrow-scope model fails on the interpretation data, and is at most borderline compatible with the production data; that the wide-scope model is able to reproduce the production data, but fails on the interpretation data; and that only the intermediate-scope model is able to fully recover the data from both conditions.

% \begin{table}[ht]
% \centering

% % \begin{tabular}{lrr}
% %   \toprule
% %   model        & production & interpretation \\ \midrule
% %   narrow       & 0.05       & 0.00 \\
% %   wide         & 0.51       & 0.00 \\
% %   intermediate & 0.49       & 0.26 \\
% %   RSA          & 0.49       & 0.52 \\
% %   \bottomrule \\
% % \end{tabular}

% \begin{tabular}{lrrrr}
%   \toprule
%   & narrow & wide & intermediate & RSA \\ \midrule
%   production & 0.05 & 0.51 & 0.49 & 0.49 \\
%   interpretation & 0.00 & 0.00 & 0.26 & 0.52 \\ \bottomrule \\
% \end{tabular}

% \caption{Sample-based estimates of Bayesian posterior predictive $p$-values for each model and each condition, based on likelihood of the observed data as test statistic.}
% \label{tab:Bppp-values}
% \end{table}
